Principles for Open Data Curation: A Case Study with the New York City 311 Service Request Data

### 2024-09-15 manuscript/manuscript.tex
connected to ML in intro

In the early 21st century, the open data movement began 
to take shape, driven by the fundamental belief that 
freely accessible data can transform both societies and 
governments. This movement champions the principles
of transparency, innovation, and public engagement. 
A landmark in this journey was the launch of the United States'
\href{https://www.data.gov}{Data.gov} portal in 2009, a pioneering
platform in making government data widely accessible. Shortly after,
the European Union followed suit, unveiling its
\href{https://data.europa.eu/euodp}{Open Data Portal} in 2012, further
cementing the movement's global reach. Furthermore, the World Bank's Open
Data initiative, initiated in 2010, stands out as a comprehensive
repository for global development data, available at
\href{https://data.worldbank.org}{World Bank Open Data}. 
These initiatives represent significant strides in democratizing data, 
in breaking barriers that once kept valuable information 
on government performance in silos. Their collective impact 
is profound, extending beyond mere data sharing to 
fostering a culture of openness that benefits individuals, 
communities, governments, and economies worldwide 
\citep{barns2016mine, wang2016adoption}.

@@

The civil and research applications of open data critically depend on
its quality, making data curation fundamental in the open data
ecosystem. Among the earliest discussions,
\citet{witt2009constructing} focused on developing data curation
profiles tailored to specific contexts, setting a precedent for
targeted data management strategies. Addressing broader challenges in
data sharing and management, \citet{borgman2012conundrum} highlighted
the complexities of research data distribution, emphasizing the need
for robust strategies. This is complemented by \citet{hart2016ten},
who outlined essential principles for effective data management,
particularly emphasizing meticulous curation practices. In
collaborative data management, \citet{beheshti2019datasynapse}
underscored the significance of cooperative environments for managing
and sharing social data effectively. This aspect gains further
relevance in \citet{mclure2014data}, which delved into the specific
practices and needs within data curation communities. The practical
implications of data curation are vividly illustrated in public health
and global challenges. \citet{cantor2018facets} demonstrated the
utility of curated open data in evaluating community health
determinants. The COVID-19 pandemic served as a real-world example,
with \citet{shankar2021data} observing the critical role of collective
data curation efforts in managing and responding to the
crisis. Collectively, these studies highlight the multifaceted nature
of data curation and emphasize its indispensable role in enhancing the
applicability and value of open data across various domains.

The contributions of this paper are twofold. First, we delve into
the specifics of data curation challenges using the NYC 311 Service
Request (SR) Data as a case study. This renowned and frequently viewed 
dataset serves as a prime example for examining key issues in data curation, 
including data validity, consistency, and curation efficiency. 
We illustrate these points with live examples drawn from our 
processing of the 311 SR data. Secondly, building upon insights 
gained from this case study, we propose a set of data curation 
principles tailored for government-released open data. These 
principles are designed to address the unique challenges 
and requirements observed in the curation of such datasets.

>>>>

In the early 21st century, the open data movement began 
to take shape, driven by the fundamental belief that 
freely accessible data can transform both societies and 
governments. This movement champions the principles
of transparency, innovation, and public engagement. 
A landmark in this journey was the launch of the United States'
\href{https://www.data.gov}{Data.gov} portal in 2009, a pioneering
platform in making government data widely accessible. Shortly after,
the European Union followed suit, unveiling its
\href{https://data.europa.eu/euodp}{Open Data Portal} in 2012, further
cementing the movement's global reach. Furthermore, the World Bank's Open
Data initiative, initiated in 2010, stands out as a comprehensive
repository for global development data, available at
\href{https://data.worldbank.org}{World Bank Open Data}. 
These initiatives represent significant strides in democratizing data, 
breaking down barriers that once kept valuable information 
on government performance in silos. Their collective impact 
extends beyond mere data sharing to fostering a culture of openness 
that benefits individuals, communities, governments, and economies worldwide 
\citep{barns2016mine, wang2016adoption}.

@@

Data curation, the process of organizing, maintaining, and ensuring
the quality of datasets, plays a crucial role in maximizing the
utility of open data. Proper curation ensures that datasets remain
consistent, accurate, and useful for diverse applications. For example, 
curating data ensures it can be leveraged effectively for machine learning 
systems, which require high-quality data to produce reliable insights
\citep{polyzotis2019data, jain2020overview}. Poor 
curation may result in issues such as missing data, formatting errors, or 
inconsistent values that can lead to biased or inaccurate outcomes
\citep{geiger2020garbage}.
This is particularly critical in domains where machine learning is applied 
to sensitive tasks, such as public health or policy
\citep{rahm2000data}.

Research into data curation has explored these challenges in depth. 
Among the earliest discussions, \citet{witt2009constructing} focused 
on developing data curation profiles tailored to specific contexts, 
setting a precedent for targeted data management strategies. 
Addressing broader challenges in data sharing and management, 
\citet{borgman2012conundrum} highlighted the complexities of 
research data distribution, emphasizing the need for robust 
strategies. This is complemented by \citet{hart2016ten}, who outlined 
essential principles for effective data management, particularly 
emphasizing meticulous curation practices. The utility of curated 
open data is vividly illustrated in public health and global challenges, 
where \citet{cantor2018facets} demonstrated the utility of curated 
data in evaluating community health determinants, and 
\citet{shankar2021data} observed its critical role during the 
COVID-19 pandemic in managing collective responses.

The contributions of this paper are twofold. First, we delve into
the specifics of data curation challenges using the NYC 311 Service
Request (SR) Data as a case study. This dataset serves as a prime 
example for examining key issues in data curation, including data 
validity, consistency, and curation efficiency. 
We illustrate these points with live examples drawn from our 
processing of the 311 SR data. Secondly, building upon insights 
gained from this case study, we propose a set of data curation 
principles tailored for government-released open data. These 
principles are designed to address the unique challenges 
and requirements observed in the curation of such datasets.
